---
title: "Merging the postal code and municipality number of Norway"
author: "Keita Abe"
date: "`r Sys.Date()`"
output: 
  minidown::mini_document:
    framework: water
    theme: light
    code_folding:
      source: hide
      output: show
      message: hide
      warning: hide
      error: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,warning = FALSE)

# library
pacman::p_load(
  dplyr,
  tidyr,
  readr,
  purrr,
  stringr,
  readxl,
  lubridate,
  knitr,
  kableExtra,
  skimr,
  patchwork
)


# load Japan map

##jp_sh = readRDS("gadm36_JPN_0_sf.rds") # 国 (country)
##jp_sh = readRDS("gadm36_JPN_1_sf.rds") # 都道府県 (prefecture)
##jp_sh = readRDS("gadm36_JPN_2_sf.rds") # 市町区村 (cities/towns)

##jp_sh2 = st_simplify(jp_sh, preserveTopology = TRUE, dTolerance = 0.1)
## pryr::object_size(jp_sh)
## pryr::object_size(jp_sh2)


# ggplot test
##ggplot(jp_sh2) + 
  ##geom_sf()
  
```



# Introduction

In Norway, there are two types of codes that represents the geographic locations: postal code and municipality number. Postal code (postnummer) is a 4 digits number, mainly used for postal address. Municipality number is also 4 digits, which represents the municipality (kommune), which is a administrative unit. The first two digits of municipality number is corresponding to the county (fylke) which is a larger administrative unit. This two digits of county number is in accordance with ISO 3166-2:NO.((New county number after 2020 has not been published in ISO3166-2 )) 

Some data includes postal codes as location information, but others use municipality number. It is convenient to have a compatible data between municipality number and postal code. The issue is that Norway experienced municipality/county mergers and the numbers are accordingly changed. 

The major mergers of counties are in two times: 2018 and 2020. On January 1st of 2018, Nord-Trøndelag and Sør-Trøndelag counties are merged as Trøndelag. 16 and 17 were assigned to these two counties, but the new county uses 50 as the county code, and the old codes are depreciated. On January 1st of 2020, there happened large merger event and the number of counties became 11 from 19. 

- Finnmark and Troms are merged as *Troms og Finnmark* and assigned the code *54*
- Hordaland and Sogn og Fjordane are merged as *Vestland* and assigned the code *46*.
- Aust-Agder and Vest-Agder are merged as *Agder* and assigned the code *42*.
- Vestfold and Telemark are merged as *Vestfold og Telemark* and assigned the code *38*.
- Hedmark and Oppland are merged as *Innlandet* and assigned the code *34*.
- Akershus, Buskerud and Østfold are merged as *Viken* and assinged the code *30*

Accordingly, the municipality code are changed. 

# The municipality code in 2019 and 2020

The data of new municipality codes after 2020 mergers are available on [Kartverket]( https://www.kartverket.no/en/kommunereform/tekniske-endringer/kommune--og-regionsendringer-2020/)


## original county and municipality data 2019-2020


```{r}
dat_20 <- readxl::read_xlsx("fylker-kommuner-2019-2020-alle.xlsx") %>%
  # change column names into English
  set_names("county_num_2019","county_name_2019","muni_num_2019","muni_name_2019",
            "county_num_2020","county_name_2020","muni_num_2020","muni_name_2020")


```




## The postal code in 2020

The data of postal code in 2020 is available on [bring.no](https://www.bring.no/radgivning/sende-noe/adressetjenester/postnummer/postnummertabeller-veiledning)

```{r}
dat_post_20 <- readxl::read_xlsx("Postnummerregister-Excel.xlsx") %>%
  # change column names into English
  set_names("post_code_2020","post_adrs_2020","muni_num_2020","muni_name_2020","post_cat")

```

category (post_cat) represnts what type of postal code is. 

- B: Both street addresses and mailboxes
- F: Multiple uses (common)
- G: Street addresses (and location addresses), ie "green mailboxes"
- P: Mailboxes
- S:Service postal code (these postal codes are not used for postal addresses)

## Origianl postal code data before 2020 (2003-2019)

```{r}
dat_post_2003_2019 <- readxl::read_xlsx("Nye, endrede og opphørte postnummer.xlsx") %>%
  mutate(Dato_kronlogisk = as.Date(Dato_kronlogisk))

dat_type <- distinct(dat_post_2003_2019) %>%
  distinct(Type_endring) %>%
  mutate(Type_change = c("Changed primary municipality", "Changed place name","Discontinued and replaced","Discontinued, not replaced","Changed category","newly created","Changed zip code"))

dat_post_2003_2019 <- dat_post_2003_2019 %>%
  # remove the change in 2020 Jan 1st because it's in dat_post_20
 # filter(Dato_kronlogisk != "2020-01-01" | Type_endring != "Endret primærkommune") %>%
  left_join(dat_type)

```

# Old and 2020 County data

As described above, county has changed by mergers in 2018 and 2020. 
The data below include which current county belonged to the previous counties. 

```{r}
dat_cnty_20 <- dat_20 %>%
  distinct(county_name_2019,county_num_2019,county_name_2020,county_num_2020) %>%
  mutate(across(contains("name"),str_to_title)) %>%
  # add 2018 change
  left_join(data.frame(
    county_num_2019 = c(50,50),
    county_name_2019 = c("Trøndelag","Trøndelag"),
    county_num_2017 = c(17,16),
    county_name_2017 = as.character(c("Nord-Trøndelag","Sør-Trøndelag"))),
    by = c("county_num_2019","county_name_2019")) %>%
  mutate(county_name_2017 = ifelse(is.na(county_name_2017),county_name_2019,as.character(county_name_2017)),
         county_num_2017 = ifelse(is.na(county_num_2017),county_num_2019,as.character(county_num_2017))) %>%
  relocate(county_num_2020, county_name_2020) %>%
  mutate(across(contains("num"),~str_pad(.x,2,pad =0)))

# write out
write_csv(dat_cnty_20, "county_2017_2020.csv")
write_csv2(dat_cnty_20, "county_2017_2020_semicolon.csv")


```

#### Data preview

```{r}
knitr::kable(head(dat_cnty_20)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```


# Old and 2020 municipality info

The code is not elegant, but the compiled data include municipality code and municipality name in each year from 2011 to 2020. 

```{r}
dat_muni_20 = dat_20 %>%
  dplyr::rename_with(~str_replace(.,"_2019","_old"), ends_with("_2019")) %>%
  mutate(year = 2019) %>%
  relocate(ends_with("_2020"), year)

  # expand the data by years 2003-2019
  ##dat_muni_expand = dat_muni_20 %>%
  ##  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020), year = 2003:2019)

# merge the expanded data. NAs for 2003-2018 for now
#dat_muni_20 = dat_muni_20 %>% right_join(dat_muni_expand) %>%
#  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,desc(year))

# make annual data of municipality info

dat_muni_03_19 = dat_post_2003_2019 %>%
  filter(Type_change == "Changed primary municipality") %>%
  mutate(year = year(Dato_kronlogisk)-1) %>%
  distinct(year,Kommunenr, Kommune,Kommunenr_etter,Kommune_etter)
  

dat_muni_20_2 = dat_muni_20 %>%
  # remove "old" data. It turns out that old data is not complete. (some old municipalities are not included)
  dplyr::select(-ends_with("old")) %>%
  # join 2019-2020 change. 
  left_join(dat_muni_03_19 %>% filter(year == 2019), by = c("year" = "year","muni_num_2020" = "Kommunenr_etter","muni_name_2020" = "Kommune_etter")) %>%
  # change names
  rename("muni_num_2019" ="Kommunenr","muni_name_2019" ="Kommune") %>%
  group_by(muni_num_2020,muni_num_2019) %>%
  fill(muni_num_2019,muni_name_2019,.direction = "up") %>%
  ungroup() %>%
  # expand the data by 2018-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019), year = 2019:2018) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_name_2019,desc(year)) %>%
  # if the muni info 2019 is NA, replace with 2020 (means no change since 2019)
  mutate(muni_name_2019 = ifelse(is.na(muni_name_2019),muni_name_2020,muni_name_2019),
         muni_num_2019 = ifelse(is.na(muni_num_2019),muni_num_2020,muni_num_2019)) %>%
  # join 2018 data using 2019 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2018), by = c("year" = "year","muni_num_2019" = "Kommunenr_etter","muni_name_2019" = "Kommune_etter")) %>%
  # change names
  rename("muni_num_2018" ="Kommunenr","muni_name_2018" ="Kommune") %>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018) %>%
  fill(muni_num_2018,muni_name_2018,.direction = "up") %>%
  ungroup() %>%
  # ===== repeat =======
  # expand the data by 2018-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019,muni_name_2018,muni_num_2018), year = 2019:2017) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_num_2018,desc(year)) %>%
 # if the muni info 2018 is NA, replace with 2020 (means no change since 2019)
  mutate(muni_name_2018 = ifelse(is.na(muni_name_2018),muni_name_2019,muni_name_2018),
         muni_num_2018 = ifelse(is.na(muni_num_2018),muni_num_2019,muni_num_2018)) %>%
  # join 2017 data using 2018 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2017), by = c("year" = "year","muni_num_2018" = "Kommunenr_etter","muni_name_2018" = "Kommune_etter"))  %>%
  # change names
  rename("muni_num_2017" ="Kommunenr","muni_name_2017" ="Kommune") %>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018) %>%
  fill(muni_num_2017,muni_name_2017,.direction = "up") %>%
  ungroup() %>%
# ===== repeat =======
  # expand the data by 2017-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019,muni_name_2018,muni_num_2018,muni_name_2017,muni_num_2017), year = 2019:2016) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_num_2018,muni_num_2017,desc(year)) %>%
 # if the muni info 2017 is NA, replace with 2018 (means no change since 2017)
  mutate(muni_name_2017 = ifelse(is.na(muni_name_2017),muni_name_2018,muni_name_2017),
         muni_num_2017 = ifelse(is.na(muni_num_2017),muni_num_2018,muni_num_2017)) %>%
  # join 2016 data using 2017 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2016), by = c("year" = "year","muni_num_2017" = "Kommunenr_etter","muni_name_2017" = "Kommune_etter"))  %>%
  # change names
  rename("muni_num_2016" ="Kommunenr","muni_name_2016" ="Kommune") %>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018,muni_num_2017) %>%
  fill(muni_num_2016,muni_name_2016,.direction = "up") %>%
  ungroup() %>%
# ===== repeat for 2015 =======
  # expand the data by 2016-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019,muni_name_2018,muni_num_2018,muni_name_2017,muni_num_2017,muni_name_2016,muni_num_2016), year = 2019:2015) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,desc(year)) %>%
 # if the muni info 2016 is NA, replace with 2017 (means no change since 2017)
  mutate(muni_name_2016 = ifelse(is.na(muni_name_2016),muni_name_2017,muni_name_2016),
         muni_num_2016 = ifelse(is.na(muni_num_2016),muni_num_2017,muni_num_2016)) %>%
  # join 2016 data using 2016 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2015), by = c("year" = "year","muni_num_2016" = "Kommunenr_etter","muni_name_2016" = "Kommune_etter"))  %>%
  # change names
  rename("muni_num_2015" ="Kommunenr","muni_name_2015" ="Kommune") %>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016) %>%
  fill(muni_num_2015,muni_name_2015,.direction = "up") %>%
  ungroup() %>%
  # ===== repeat for 2014 =======
  # expand the data by 2015-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019,muni_name_2018,muni_num_2018,muni_name_2017,muni_num_2017,muni_name_2016,muni_num_2016,muni_name_2015,muni_num_2015), year = 2019:2014) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,desc(year)) %>%
 # if the muni info 2015 is NA, replace with 2016 (means no change since 2017)
  mutate(muni_name_2015 = ifelse(is.na(muni_name_2015),muni_name_2016,muni_name_2015),
         muni_num_2015 = ifelse(is.na(muni_num_2015),muni_num_2016,muni_num_2015)) %>%
  # join 2016 data using 2016 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2014), by = c("year" = "year","muni_num_2015" = "Kommunenr_etter","muni_name_2015" = "Kommune_etter"))  %>%
  # change names
  rename("muni_num_2014" ="Kommunenr","muni_name_2014" ="Kommune") %>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015) %>%
  fill(muni_num_2014,muni_name_2014,.direction = "up") %>%
  ungroup() %>%
  # ===== repeat for 2013 =======
  # expand the data by 2014-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019,muni_name_2018,muni_num_2018,muni_name_2017,muni_num_2017,muni_name_2016,muni_num_2016,muni_name_2015,muni_num_2015,muni_name_2014,muni_num_2014), year = 2019:2013) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,muni_num_2014,desc(year)) %>%
 # if the muni info 2014 is NA, replace with 2015 (means no change since 2015)
  mutate(muni_name_2014 = ifelse(is.na(muni_name_2014),muni_name_2015,muni_name_2014),
         muni_num_2014 = ifelse(is.na(muni_num_2014),muni_num_2015,muni_num_2014)) %>%
  # join 2013 data using 2014 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2013), by = c("year" = "year","muni_num_2014" = "Kommunenr_etter","muni_name_2014" = "Kommune_etter"))  %>%
  # change names to 2013
  rename("muni_num_2013" ="Kommunenr","muni_name_2013" ="Kommune")%>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,muni_num_2014) %>%
  fill(muni_num_2013,muni_name_2013,.direction = "up") %>%
  ungroup() %>%
  # ===== repeat for 2012 =======
  # expand the data by 2013-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019,muni_name_2018,muni_num_2018,muni_name_2017,muni_num_2017,muni_name_2016,muni_num_2016,muni_name_2015,muni_num_2015,muni_name_2014,muni_num_2014,muni_name_2013,muni_num_2013), year = 2019:2012) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,muni_num_2014,muni_num_2013,desc(year)) %>%
 # if the muni info 2013 is NA, replace with 2014 (means no change since 2014)
  mutate(muni_name_2013 = ifelse(is.na(muni_name_2013),muni_name_2014,muni_name_2013),
         muni_num_2013 = ifelse(is.na(muni_num_2013),muni_num_2014,muni_num_2013)) %>%
  # join 2012 data using 2013 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2012), by = c("year" = "year","muni_num_2013" = "Kommunenr_etter","muni_name_2013" = "Kommune_etter"))  %>%
  # change names to 2012
  rename("muni_num_2012" ="Kommunenr","muni_name_2012" ="Kommune")%>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,muni_num_2014,muni_num_2013) %>%
  fill(muni_num_2012,muni_name_2012,.direction = "up") %>%
  ungroup() %>%
  # ===== repeat for 2011 =======
  # expand the data by 2012-2019
  expand(nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_name_2019,muni_num_2019,muni_name_2018,muni_num_2018,muni_name_2017,muni_num_2017,muni_name_2016,muni_num_2016,muni_name_2015,muni_num_2015,muni_name_2014,muni_num_2014,muni_name_2013,muni_num_2013,muni_name_2012,muni_num_2012), year = 2019:2011) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,muni_num_2014,muni_num_2013,muni_num_2012,desc(year)) %>%
 # if the muni info 2012 is NA, replace with 2013 (means no change since 2014)
  mutate(muni_name_2012 = ifelse(is.na(muni_name_2012),muni_name_2013,muni_name_2012),
         muni_num_2012 = ifelse(is.na(muni_num_2012),muni_num_2013,muni_num_2012)) %>%
  # join 2011 data using 2012 names. 
  left_join(dat_muni_03_19 %>% filter(year == 2011), by = c("year" = "year","muni_num_2012" = "Kommunenr_etter","muni_name_2012" = "Kommune_etter"))  %>%
  # change names to 2011
  rename("muni_num_2011" ="Kommunenr","muni_name_2011" ="Kommune") %>%
  group_by(muni_num_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,muni_num_2014,muni_num_2013,muni_num_2012) %>%
  fill(muni_num_2011,muni_name_2011,.direction = "up") %>%
  ungroup() %>%
 # ==== End: No municipality change before 2011 in the data ====
# if the muni info 2011 is NA, replace with 2013 (means no change since 2014)
  mutate(muni_name_2011 = ifelse(is.na(muni_name_2011),muni_name_2012,muni_name_2011),
         muni_num_2011 = ifelse(is.na(muni_num_2011),muni_num_2012,muni_num_2011)) %>%
  # remove overlapping data* names are overlapping by fill to join the previous year data. We don't need it now.
  distinct(muni_num_2020,muni_num_2019,muni_num_2018,muni_num_2017,muni_num_2016,muni_num_2015,muni_num_2014,muni_num_2013,muni_num_2012,muni_num_2011,.keep_all = TRUE) %>%
  dplyr::select(-year)

# write out
##write_csv(dat_muni_20_2, "municipality_2011_2020.csv")
##write_csv2(dat_muni_20_2, "municipality_2011_2020_semicolon.csv")

# long format: each row is 2020 info x each year of old info.
dat_muni_20_2_long = dat_muni_20_2 %>%
  pivot_longer(-c(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020),
               names_to = c(".value","year"),
               names_pattern = "(.*_.*)_(.*)") %>%
  rename("muni_name_old" = "muni_name", "muni_num_old" = "muni_num") %>% 
  mutate(year = as.numeric(year),
         county_num_2020 = as.character(str_pad(county_num_2020,2,pad = "0")),
         muni_num_2020 = as.character(str_pad(muni_num_2020,4,pad = "0")),
         muni_num_old = as.character(str_pad(muni_num_old,4,pad = "0")))

# extend back to 2002
dat_muni_20_3_long = dat_muni_20_2_long %>%
  right_join(expand(dat_muni_20_2_long,nesting(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020), year = 2002:2019), by = c("county_num_2020","county_name_2020","muni_num_2020","muni_name_2020","year")) %>%
  arrange(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020,year) %>%
  group_by(county_num_2020,county_name_2020,muni_num_2020,muni_name_2020) %>%
  fill(muni_num_old,muni_name_old,.direction = "up") %>%
  ungroup()

# write out
write_csv(dat_muni_20_3_long, "municipality_2011_2020_long.csv")
write_csv2(dat_muni_20_3_long, "municipality_2011_2020_long_semicolon.csv")


```


#### Data preview

```{r}
knitr::kable(head(dat_muni_20_3_long)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```


# Old and 2020 postal code data

The postal code data in 2020 and corresponding old information back to 2002.
Annual data is determined as of Decemebr 31st of each year.
For example, if a change is made in Octber 1st in 2013, 2013 data reflects the change. 


```{r}
# the main data of post address is dat_post_20

# old data 2003-2019
dat_post_03_19 <- dat_post_2003_2019 %>%
  mutate(year = year(Dato_kronlogisk)-1) %>%
  # because newly created post address is "after change" address.
  mutate(Postnummer_etter = ifelse(Type_endring == "Nyopprettet",Postnummer,Postnummer_etter),
         Poststedsnavn_etter = ifelse(Type_endring == "Nyopprettet",Poststedsnavn,Poststedsnavn_etter),
         Kommunenr_etter = ifelse(Type_endring == "Nyopprettet",Kommunenr,Kommunenr_etter),
         Kommune_etter = ifelse(Type_endring == "Nyopprettet",Kommune,Kommune_etter),
         Kategori_etter = ifelse(Type_endring == "Nyopprettet",Kategori,Kategori_etter)) %>%
  # change the numeber to character and leading 0 (e.g. 301 is 0301)
  mutate(Postnummer_etter = as.character(str_pad(Postnummer_etter,4,pad = "0")),
         Postnummer = as.character(str_pad(Postnummer,4,pad = "0")))

# extract discontinued address and attach to the main data
dat_discont = dat_post_03_19 %>% filter(Type_endring == "Opphørt, ikke erstattet") %>%
  dplyr::select(Dato_kronlogisk,Type_endring,Postnummer,Poststedsnavn,Kommunenr,Kommune,Kategori,Type_change,year)

dat_discont2 = dat_discont %>%
  # expand to fill the old info.
  right_join(expand(dat_discont,nesting(Postnummer, Poststedsnavn), year = 2002:2019)) %>%
  arrange(Postnummer,year) %>%
  group_by(Postnummer,Poststedsnavn) %>%
  fill(Kommunenr,Kommune,Kategori, .direction = "up") %>%
  ungroup() %>%
  # remove "new" data (because discontinued)
  filter(!is.na(Dato_kronlogisk))

# merge the changed log based on "after change" address as address in 2020
dat_post_20_2 <- dat_post_20 %>%
    expand(nesting(post_code_2020,post_adrs_2020,muni_num_2020,muni_name_2020,post_cat), year = 2002:2019) %>%
  left_join(dat_post_03_19, by = c("post_code_2020" = "Postnummer_etter", "post_adrs_2020" = "Poststedsnavn_etter", "year" = "year")) %>%
  arrange(post_code_2020,post_adrs_2020,year) %>%
  group_by(post_code_2020, post_adrs_2020) %>%
  # fill the old information
  # do not fill "Type_change" in order to indicate when it was changed. 
  fill(Postnummer,Poststedsnavn,Kommunenr,Kommune,Kategori,Kommunenr_etter,Kommune_etter,Kategori_etter, .direction = "up") %>%
  ungroup() %>%
  bind_rows(dat_discont2) 

# check: not attachd to the main data
temp = anti_join(dat_post_03_19,dat_post_20_2) %>%
  filter(Type_endring != "Opphørt, ikke erstattet")

# why are they remaining? Because the "after change" address does not exist in 2020.
# "after change" was also changed later.
# -> merge to the old address info

# idea: make the df that merges un-attached rows and 2020 info based on "before change" data in 2020
# then bind_rows with the main data
temp2 = temp %>%
  left_join(dat_post_20_2,by = c("Postnummer_etter" = "Postnummer","Poststedsnavn_etter" ="Poststedsnavn", "year" = "year")) %>%
  dplyr::select(-ends_with(".y")) %>%
  rename_with(~str_replace(.,".x",""),ends_with(".x"))

# merge (bind) unattached data
dat_post_20_3 <- dat_post_20_2 %>% 
  bind_rows(temp2) %>%
  # fill old information if NA. replaced with the current info
  mutate(Postnummer = ifelse(is.na(Postnummer),post_code_2020,Postnummer),
         Poststedsnavn = ifelse(is.na(Poststedsnavn),post_adrs_2020,Poststedsnavn),
         Kategori = ifelse(is.na(Kategori),post_cat,Kategori)) %>%
  dplyr::select(post_code_2020,post_adrs_2020,muni_num_2020,muni_name_2020,post_cat,
                year, Postnummer, Poststedsnavn, Kommunenr, Kommune, Kategori, Type_change, Dato_kronlogisk) %>%
  # rename to English
  rename("post_code_old" = "Postnummer", "post_adrs_old" = "Poststedsnavn", "muni_code_old" = "Kommunenr", "muni_name_old" =  "Kommune",  "post_cat_old" = "Kategori", "Date_change" = "Dato_kronlogisk") %>%
  arrange(post_code_2020,post_adrs_2020,year) %>%
  # discontinued addresses, assigned 9999 postal code
  mutate(post_code_2020 = ifelse(is.na(post_code_2020),"9999",post_code_2020),
         post_adrs_2020 = ifelse(is.na(post_adrs_2020),"Discontinued",post_adrs_2020)) %>%
  # merge old municipalities info
  left_join(dat_muni_20_3_long, by = c("year", "muni_num_2020")) %>%
  mutate(muni_code_old = ifelse(is.na(muni_code_old), muni_num_old,muni_code_old),
         muni_name_old.x = ifelse(is.na(muni_name_old.x), muni_name_old.y,muni_name_old.x)) %>%
  dplyr::select(-ends_with(".y"),-muni_num_old) %>%
  rename_with(~str_replace(.,".x",""),ends_with(".x")) %>%
  relocate(county_num_2020,county_name_2020, .after = muni_name_2020) %>%
  # merge old county names and code
  mutate(muni_code_old = str_pad(muni_code_old,4,pad = 0)) %>%
  # change all title letters 
  mutate(across(contains("name"),str_to_title)) %>%
  left_join(dat_cnty_20,by = c("county_num_2020","county_name_2020")) %>%
  mutate(county_num_old = ifelse(year >= 2018, county_num_2019,county_num_2017),
         county_name_old = ifelse(year >= 2018, county_name_2019,county_name_2017)) %>%
  # remove the 2019 and 2017 county info
  dplyr::select(-ends_with("_2019"),-ends_with("_2017")) %>%
  relocate(county_num_old, county_name_old, .after = muni_name_old)

# write out
write_csv(dat_post_20_3,"postal_code_2002_2020_long.csv")
write_csv2(dat_post_20_3,"postal_code_2002_2020_long_semicolon.csv")
```

#### Data preview

```{r}
knitr::kable(head(dat_post_20_3))  %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```


Task left: Old Municipality/county info in the discontinued post addresses are missing. 
